---
title: "Color Vision"
author: "Emmanuel Maduneme"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## Color Vision

Use the [colorbrewer](https://colorbrewer2.org/#type=sequential&scheme=BuGn&n=3) to select color palettes that are colorblind friendly.

Another site for colorblind friendly: [coolors](https://coolors.co/).

```{r}
## Use this package to for color palettes
p_load(colorspace, scales)

## to call the palletes
hcl_palettes(plot = T)

## Scales pkg helps you get the hexcodes
q4 <- colorspace::qualitative_hcl(4, palette = "Dark 3")

q4
show_col(q4)

## You can check those with CVD or CB can see the color
show_col(deutan(q4))

```


Use `colorblindr` to check how plot shows for different cvd. It can help pick after plot has been made. 
```{r}

remotes::install_github("wilkelab/cowplot")
install.packages("colorspace", repos = "http://R-Forge.R-project.org")


remotes::install_github("clauswilke/colorblindr")


```


After getting your colorbrewer palettes, create a vector of selected hexcodes and then use the ggplot function: `scale_fill_manual(values= vector)`, 
